---
title: "Ex7"
author: "Guanting Wei"
date: "Oct.24.2018"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# 1. MCMC With Gibbs Sampling
We have parameters below.
$$\mu_1 \sim N(0,10^2)$$
$$\mu_2 \sim N(0,10^2)$$
$$\frac{1}{\sigma_1^2}\sim \Gamma(a,b)$$
$$\frac{1}{\sigma_2^2}\sim \Gamma(a,b)$$
$$a=0.5, \ b=10, \lambda=\frac{1}{b}=\frac{1}{10}$$
$$\sigma_1^2\sim IG(0.5,10)$$
$$\sigma_2^2\sim IG(0.5,10)$$
$$\delta\sim U(0,1)$$

The density function of X is a mixture normal distribution.
$$f(x)=\delta \frac{1}{\sqrt{2\pi}\sigma_1}e^{-\frac{(x-\mu_1)^2}{2\sigma_1^2}}+(1-\delta)\frac{1}{\sqrt{2\pi}\sigma_2}e^{-\frac{(x-\mu_2)^2}{2\sigma_2^2}}$$
The Likelihood function is as follow. 
$$L(x;\delta,\mu_1,\sigma_1^2,\mu_2,\sigma_2^2)=\prod_{i=1}^{n} (\delta \frac{1}{\sqrt{2\pi}\sigma_1}e^{-\frac{(x-\mu_1)^2}{2\sigma_1^2}}+(1-\delta)\frac{1}{\sqrt{2\pi}\sigma_2}e^{-\frac{(x-\mu_2)^2}{2\sigma_2^2}})$$
The joint posterior probability is as follow.
\[
\begin{aligned}
P(\delta,\mu_1,\sigma_1^2,\mu_2,\sigma_2^2|x)&\propto L(x;\delta,\mu_1,\sigma_1^2,\mu_2,\sigma_2^2)*f_1(\mu_1)*f_2(\mu_2)*f_3(\sigma_1^2)*f_4(\sigma_2^2)
\\&=\prod_{i=1}^{n} (\delta \frac{1}{\sqrt{2\pi}\sigma_1}e^{-\frac{(x-\mu_1)^2}{2\sigma_1^2}}+(1-\delta)\frac{1}{\sqrt{2\pi}\sigma_2}e^{-\frac{(x-\mu_2)^2}{2\sigma_2^2}})\\&\times 1\times\frac{1}{10\sqrt{2\pi}}e^{-\frac{\mu_1^2}{2\times 10^2}}\times\frac{1}{\sqrt{10}\sqrt{2\pi}}\frac{1}{(\sigma_1^2)^\frac{3}{2}}e^{-\frac{1}{10\sigma_1^2}}\times\frac{1}{10\sqrt{2\pi}}e^{-\frac{\mu_2^2}{2\times 10^2}}\times\frac{1}{\sqrt{10}\sqrt{2\pi}}\frac{1}{(\sigma_2^2)^\frac{3}{2}}e^{-\frac{1}{10\sigma_2^2}}
\end{aligned}
\]

We need to get
$$P(\delta|x,\mu_1,\sigma_1^2,\mu_2,\sigma_2^2),P(\mu_1|x,\delta,\sigma_1^2,\mu_2,\sigma_2^2),P(\sigma_1^2|x,\delta,\mu_1,\mu_2,\sigma_2^2),P(\mu_2|x,\delta,\mu_1,\sigma_1^2,\sigma_2^2),P(\sigma_2^2|x,\delta,\mu_1,\sigma_1^2,\mu_2)$$

$\\$
$\\$
$\\$
$\\$

# 2. Code And Graph
```{r}
library(actuar)
delta <- 0.7 # true value to be estimated based on the data
n <- 100
set.seed(123)
u <- rbinom(n, prob = delta, size = 1)
x <- rnorm(n, ifelse(u == 1, 7, 10), 0.5)

loglikelihood=function(delta,mu1,mu2,var1,var2){
  sum(log(delta*dnorm(x,mu1,sqrt(var1))+(1-delta)*dnorm(x,mu2,sqrt(var2))))
}
prior=function(delta,mu1,mu2,var1,var2){
  delta.prior=1
  mu1.prior=dnorm(mu1,0,10)
  mu2.prior=dnorm(mu2,0,10)
  var1.prior=dinvgamma(var1,0.5,10)
  var2.prior=dinvgamma(var2,0.5,10)
  log(mu1.prior*mu2.prior*var1.prior*var2.prior)
}
all=function(delta,mu1,mu2,var1,var2){
  loglikelihood(delta,mu1,mu2,var1,var2)+prior(delta,mu1,mu2,var1,var2)
}
library(HI)
mcmcgibbs=function(delta.start,mu1.start,mu2.start,var1.start,var2.start,count){
  delta.new=rep(0,count)
  mu1.new=rep(0,count)
  mu2.new=rep(0,count)
  var1.new=rep(0,count)
  var2.new=rep(0,count)
  for(i in 1:count){
    delta.new[i]=arms(delta.start,all,function(x,mu1,mu2,var1,var2)(x>0)*(x<1),1,
                      mu1=mu1.start,mu2=mu2.start,var1=var1.start,var2=var2.start)
    mu1.new[i]=arms(mu1.start,all,function(delta,x,mu2,var1,var2)(x>-50)*(x<50),1,
                    delta=delta.start,mu2=mu2.start,var1=var1.start,var2=var2.start)
    mu2.new[i]=arms(mu2.start,all,function(delta,mu1,x,var1,var2)(x>-50)*(x<50),1,
                    mu1=mu1.start,delta=delta.start,var1=var1.start,var2=var2.start)
    var1.new[i]=arms(var1.start,all,function(delta,mu1,mu2,x,var2)(x>0.00001)*(x<100),1,
                     mu1=mu1.start,mu2=mu2.start,delta=delta.start,var2=var2.start)
    var2.new[i]=arms(var2.start,all,function(delta,mu1,mu2,var1,x)(x>0.00001)*(x<100),1,
                     mu1=mu1.start,mu2=mu2.start,var1=var1.start,delta=delta.start)
    delta.start=delta.new[i]
    mu1.start=mu1.new[i]
    mu2.start=mu2.new[i]
    var1.start=var1.new[i]
    var2.start=var2.new[i]
  }
  rbind(delta.new,mu1.new,mu2.new,var1.new,var2.new)
}
```

With initial value$\delta^{(0)}=0.5,\mu_1^{(0)}=0,{\sigma_1^2}^{(0)}=0,\mu_2^{(0)}=10,{\sigma_2^2}^{(0)}=10$, we run the 5000 times and throw away the first 2500.

```{r}
data=mcmcgibbs(0.5,0,0,10,10,5000)[,-(1:2500)]

```

```{r}
library(ggplot2)

ggplot(data.frame(x=data[1,]),aes(x=x))+
xlab(~delta)+
ylab("density")+
ggtitle(~delta)+
geom_histogram(aes(y=..density..),fill="gray", colour="black")
plot(ts(data[1,]),xlab="Time",ylab=~delta)

ggplot(data.frame(x=data[2,]),aes(x=x))+
xlab(~mu[1])+
ylab("density")+
ggtitle(~mu[1])+
geom_histogram(aes(y=..density..),fill="gray", colour="black")
plot(ts(data[2,]),xlab="Time",ylab=~mu[1])

ggplot(data.frame(x=data[3,]),aes(x=x))+
xlab(~mu[2])+
ylab("density")+
ggtitle(~mu[2])+
geom_histogram(aes(y=..density..),fill="gray", colour="black")
plot(ts(data[3,]),xlab="Time",ylab=~mu[2])

ggplot(data.frame(x=data[4,]),aes(x=x))+
xlab(~sigma[1]^2)+
ylab("density")+
ggtitle(~sigma[1]^2)+
geom_histogram(aes(y=..density..),fill="gray", colour="black")
plot(ts(data[4,]),xlab="Time",ylab=~sigma[1]^2)

ggplot(data.frame(x=data[5,]),aes(x=x))+
xlab(~sigma[2]^2)+
ylab("density")+
ggtitle(~sigma[2]^2)+
geom_histogram(aes(y=..density..),fill="gray", colour="black")
plot(ts(data[5,]),xlab="Time",ylab=~sigma[1]^2)
```

